Think deep for reef resilience

### 2025-08-06 paper/paper_draft.qmd
Comments for Future DHW section
Still thinking on section header as there's a slight repetition now.

Need to double check once I have plot generation working but I think the y-axis is wrong for the two figures maybe?

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025
to 1. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef depth and the
number of years cover exceeds the carbonate budget threshold. Point colour represents the
median reef depth level in metres. Displays results from median timeseries using
ACCESS-ESM1-5.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/depth_carbonate_scatter.png){#fig-access-esm1-5-depth-carbonate width=14.82cm}

## Global Climate Model future trajectories

Each GCM shows similar relationships between reef clusters and contextual analysis factors,
thus further comparisons of these clusters for each analysis factor are only shown for
ACCESS-ESM1-5 (intermediate ECS scenario). Results for the additional GCMs are displayed in
the supplementary material. However, there are substantial differences in the longevity of
reef resilience and maintenance of coral cover post 2030 between the GCMs, thus coral cover
timeseries are displayed for each GCM (displayed clustered at management-area scale to
preserve coarse spatial differences in coral trajectories) (@fig-gcm-cover-trajectories).

NorESM2-MM has the lowest Equilibrium Climate Sensitivity (2.5°C) and reefs exhibit
continuing recovery and decline dynamics up to 2075 under this possible climate future.
While recovery phases are predominantly seen in the ‘high’ cover cluster, the ‘medium’
cluster, and to a lesser degree the ‘low’ cluster, also exhibit periods of recovery and
increasing proportions of coral cover up to 2045. Under NorESM2-MM the Far Northern
management area does not see coral cover maintenance as high as the three remaining
management regions. Under the second lowest ECS projection (GFDL-CM4, 2.9°C) all clusters in
each management area exhibit largely declining future trajectories with periods of recovery
from 2030 to 2040. ACCESS-ESM1-5 has an intermediate ECS value of 3.87°C which results in
intermediate levels of coral cover maintenance (between those of lower and higher ECS GCMs).
In Townsville/Whitsunday and Far Northern management areas reefs in the high cluster are
able to maintain coral cover levels around 20% to 2050 under ACCESS-ESM1-5. Under ACCESS-CM2
(4.72°C) and EC-Earth3-Veg (4.31°C) GCMs cover trajectories are largely declining in all
management areas.

Under each GCM, in all management areas, low, medium and high cluster reefs have low
differences between DHW trajectories (@fig-gcm-dhw-trajectories). The highest peak DHW
levels occur under GFDL-CM4 between 2045 and 2050, while greater increases in overall DHW
timeseries occur under ACCESS-CM2.

![Reef area-relative coral cover timeseries clustered within each management area, for each
GCM analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover
between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with
intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’
clusters (blue) are clusters with the highest median area-relative cover between 2030 and
2040 within each bioregion. Solid lines represent the cluster median timeseries and bands
represent the 2.5 to 97.5 percentile confidence intervals. Each management area is
represented five times (for each GCM), with each GCM displayed
horizontally.](../outputs/ADRIA_results/HighResCoralStress/figs/GCM_cover_timeseries_plot.png){#fig-gcm-cover-trajectories
width=14.82cm}

![Reef Degree Heating Week timeseries clustered within each management area, for each GCM
analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover
between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with
intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’
clusters (blue) are clusters with the highest median area-relative cover between 2030 and
2040 within each bioregion. Solid lines represent the cluster median timeseries and bands
represent the 2.5 to 97.5 percentile confidence intervals. Each management area is
represented five times (for each GCM), with each GCM displayed
horizontally.](../outputs/ADRIA_results/HighResCoralStress/figs/GCM_dhw_timeseries_plot.png){#fig-gcm-dhw-trajectories
width=14.82cm}

>>>>

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025
to 1. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef depth and the
number of years cover exceeds the carbonate budget threshold. Point colour represents the
median reef depth level in metres. Displays results from median timeseries using
ACCESS-ESM1-5.](../outputs/ADRIA_results/HighResCoralStress/figs/ACCESS-ESM1-5/depth_carbonate_scatter.png){#fig-access-esm1-5-depth-carbonate
width="14.82cm"}

## Future DHW trajectories

Each Global Climate Model (GCM) shows similar relationships between reef clusters and
contextual analysis factors, thus further comparisons of these clusters for each analysis
factor are only shown for ACCESS-ESM1-5 (an intermediate Equilibrium Climate Sensitivity
scenario). Results for the additional GCMs are displayed in the supplementary material.
However, there are substantial differences in the longevity of reef resilience and
maintenance of coral cover post 2030 between the GCMs, thus coral cover timeseries are
displayed for each GCM (displayed clustered at management-area scale to preserve coarse
spatial differences in coral trajectories) (@fig-gcm-cover-trajectories).

NorESM2-MM has the lowest Equilibrium Climate Sensitivity (2.5°C) and reefs exhibit
continuing recovery and decline dynamics up to 2075 under this possible climate future.
Recovery phases are predominantly seen in the ‘high’ cover cluster. The ‘medium’ cluster,
and to a lesser degree the ‘low’ cluster, also exhibit periods of recovery and increasing
proportions of coral cover up to 2045. Under NorESM2-MM the Far Northern management area
does not see coral cover maintenance as high as the other management regions \[**TI:**
**Quantify this! How did you determine this and how large/small are the difference in coral
cover maintenance?**\]. Under the second lowest ECS projection (GFDL-CM4, 2.9°C) all
clusters in each management area exhibit largely declining future trajectories with periods
of recovery from 2030 to 2040. ACCESS-ESM1-5 has an intermediate ECS value of 3.87°C which
results in intermediate levels of coral cover maintenance (between those of lower and higher
ECS GCMs). In Townsville-Whitsunday and Far Northern management areas, reefs in the high
cluster are able to maintain coral cover levels around 20% to 2050 under ACCESS-ESM1-5.
Under ACCESS-CM2 (4.72°C) and EC-Earth3-Veg (4.31°C) GCMs cover trajectories are largely
declining in all management areas.

Under each GCM, in all management areas, low, medium and high cluster reefs have low
differences between DHW trajectories (@fig-gcm-dhw-trajectories) \[**TI:** **I don't
understand this sentence - what is meant by reefs have low differences?**\]. The highest
peak DHW levels occur under GFDL-CM4 between 2045 and 2050, while greater increases in
overall DHW timeseries occur under ACCESS-CM2.

![Reef Degree Heating Week timeseries clustered within each management area, for each GCM
analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover
between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with
intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’
clusters (blue) are clusters with the highest median area-relative cover between 2030 and
2040 within each bioregion. Solid lines represent the cluster median timeseries and bands
represent the 2.5 to 97.5 percentile confidence intervals. Each management area is
represented five times (for each GCM), with each GCM displayed
horizontally.](../outputs/ADRIA_results/HighResCoralStress/figs/GCM_dhw_timeseries_plot.png){#fig-gcm-dhw-trajectories
width="14.82cm"}

![Reef area-relative coral cover timeseries clustered within each management area, for each
GCM analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover
between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with
intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’
clusters (blue) are clusters with the highest median area-relative cover between 2030 and
2040 within each bioregion. Solid lines represent the cluster median timeseries and bands
represent the 2.5 to 97.5 percentile confidence intervals. Each management area is
represented five times (for each GCM), with each GCM displayed
horizontally.](../outputs/ADRIA_results/HighResCoralStress/figs/GCM_cover_timeseries_plot.png){#fig-gcm-cover-trajectories
width="14.82cm"}
